---
title: "Mastering Software Development in R Capstone Project"
author: "Carlos Silva"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
        
```{r, echo = FALSE, include = FALSE}
library(RCapstone)
library(dplyr)
library(ggplot2)
library(readr)
library(grid)
library(lubridate)
library(leaflet)
library(scales)
```

Overview of the `RCapstone` R package created for visualizing U.S. National Oceanographic and Atmospheric Administation's (NOAA)  dataset of significant earthquakes. The data examined is from [NOAA database](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1).

## Installation

You can install RCapstone from github with:

```
library(devtools)
install_github("CarlosSilva34/RCapstone")
library(RCapstone)
```

## Package functions

Exported functions available:

- `eq_clean_data()`
- `eq_location_clean()`
- `geom_timeline()`
- `geom_timeline_label()`
- `theme_timeline()`
- `eq_create_label()`
- `eq_map()`


## Read and clean data

The first function cleans the data for the visualization. It creates a DATE column in `Date` format, transforms latitude and longitude to numeric format and trims country from LOCATION_NAME.

```{r eq_read, message = FALSE} 
filename <- system.file("extdata/signif.txt", package = "RCapstone")
data <- readr::read_delim(filename, delim = "\t")
eq_clean_data(data)
```

## Timeline of earthquakes in US and China after 2000

These functions use `ggplot2` package to visualize earthquake timeline. The`geom_timeline()` geom requires clean data. The aesthetics are `x` with dates, optional `y` for grouping by country,`size` and `color`. The `geom_timeline_label()` function requires `label` aesthetic. The `theme_timeline()` geom was added for better visualization.

```{r eq_timeline_graph, fig.width = 7, fig.height = 4}
data %>% eq_clean_data() %>%
filter(COUNTRY %in% c("USA", "CHINA"), YEAR >= 2000) %>%
ggplot(aes(x = DATE,
y = COUNTRY,
color = as.numeric(TOTAL_DEATHS),
size = as.numeric(EQ_PRIMARY)
)) +
geom_timeline() +
geom_timeline_label(aes(label = LOCATION_NAME), n_max = 5) +
theme_timeline() +
labs(size = "Richter scale value", color = "# deaths")
```

## Map of earthquakes in Mexico

These functions use `leaflet` package to visualize earthquakes on a map using `eq_map()` function. Optional annotations can be created using `eq_create_label()` function. This creates an interactive map where user can click on individual points to get details.


```{r eq_map_Mexico, fig.width = 7, fig.height = 4}
data %>% 
eq_clean_data() %>% 
dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
dplyr::mutate(popup_text = eq_create_label(.)) %>% 
eq_map(annot_col = "popup_text")
